---
title: "MoneyBball"
author: Austin Daines
date: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(dplyr)
library(tidyverse)
library(DT)
library(gapminder)
```

This project is inspired by <i>Moneyball</i>: The Art of Winning an Unfair Game, a book by Michael Lewis, published in 2003, about the Oakland Athletics baseball team and its general manager Billy Beane. Its focus is the team's analytical, evidence-based, sabermetric approach to assembling a competitive baseball team despite Oakland's small budget 
([Wiki source](https://en.wikipedia.org/wiki/Moneyball)). 

<i>In 2001 the Oakland Athletics team salary was only <b>34 million dollars</b> compared to the league leading <b>112 million dollar</b> New York Yankees. </i>

The discrepancy in team salary is not as drastic in the NBA. 

Here I took a web scraping function from [mbjoseph](https://rdrr.io/github/mbjoseph/bbr/src/R/zzz.R) to scrape NBA team salaries. 

+ Web scraping hoopshype.com/salaries...

```{r, echo=FALSE}
# -----------------------------------------------------------------------------
# maybe_as_numeric
#
# https://rdrr.io/github/mbjoseph/bbr/src/R/zzz.R

maybe_as_numeric <- function(x) {
  # tries to make numeric columns numeric (from char)
  numeric_x <- suppressWarnings(as.numeric(x))
  if (!all(is.na(numeric_x))) x <- numeric_x
  x
}

# -----------------------------------------------------------------------------
# empty_string_to_na
#

empty_string_to_na <- function(x) {
  # sometimes (especially old datasets), empty values are ""
  if (class(x) == "character") {
    res <- ifelse(x == "", NA, x)
  } else {
    res <- x
  }
  res
}

# -----------------------------------------------------------------------------
# clean_colnames
#

clean_colnames <- function(df) {
  # clean up column names for a data frame
  stopifnot(is.data.frame(df))
  df <- df[!(names(df) == "Rk")] # remove "Rank" column
  names(df) <- gsub("\\.", "_pct", names(df))
  names(df) <- gsub("X2", "two_", names(df))
  names(df) <- gsub("X3", "three_", names(df))
  names(df) <- tolower(names(df))
  df
}

# -----------------------------------------------------------------------------
# get_hoops_salary
#

get_hoops_salary <- function(url) {
  html <- xml2::read_html(url)
  node <- rvest::html_node(html, "table")
  table <- rvest::html_table(node, header = TRUE)
  converted <- lapply(table, maybe_as_numeric)
  converted <- lapply(converted, empty_string_to_na)
  df <- as.data.frame(converted, stringsAsFactors = FALSE)

  # select only the columns we need: Team name and X2020.21 salary
  df <- df %>% dplyr::select(2:3) 
  
  # convert salary from chr with $ and ,'s to numeric
  df$X2020.21 <- as.numeric(gsub('[$,]', '', df$X2020.21))
  
  return(df)
}

# ===========================================================================
# web scrape hoopshype.com for team salaries

#print("Web scraping hoopshype.com/salaries...")
df_hoops_ts <- get_hoops_salary("https://hoopshype.com/salaries/")
ts_dim <- dim(df_hoops_ts)
paste0("Loaded ", ts_dim[1], " team salaries.")

```

Here are the (6) largest team salaries:

```{r, echo=FALSE}
head(df_hoops_ts)
```

#### Boxplot of team salary: 

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.align='center', fig.width=4, fig.height=6}

ggplot(df_hoops_ts, aes(y=X2020.21/1000000)) +
  geom_boxplot(aes()) +
  theme_minimal() +
  theme(axis.text.x =element_blank()) +
  ylab("20-21 NBA Team Salaries ($M)")
  
```

In 2001 the Oakland A's paid about 1/2 the median MLB team salary. 
With this in mind we will attempt to build a competitive team with 1/2 the median NBA team salary. Here we see the median team salary in the NBA is roughly $130 million. That leaves us with 65 million dollars to work with. 

### Hypothesis

Similar to baseball, we can create a competitive team with a salary of $65M.

#### What constitutes a competitive team? 

Well, that's not a simple question to answer. We need a team with players who contribute both offensively and defensively. First, we will analyze plus-minus statistics

##### Plus-Minus 

+ [Plus-Minus](https://www.nbastuffer.com/analytics101/plus-minus/), a.k.a. +/-, simply keeps track of the net changes in the score when a given player is either on or off the court.

##### Real Plus-Minus 


+ [Real Plus-Minus](https://www.nbastuffer.com/analytics101/real-plus-minus-rpm/) can be broken down into offensive and defensive metrics:

  + Offensive Real Plus-Minus: (ORPM): Player’s average impact on his team’s offensive performance, by the points scored per 100 offensive possessions.

  + Defensive Real Plus-Minus: (DRPM): Player’s average impact on his team’s defensive performance, by the points allowed per 100 offensive possessions.


##### RPM Wins 

+ [RPM Wins](http://www.espn.com/nba/statistics/rpm#:~:text=Glossary&text=WINS%3A%20%22RPM%20Wins%22%20provide,his%20number%20of%20possessions%20played.) provide an estimate of the number of wins each player has contributed to his team's win total on the season. RPM Wins include the player's Real Plus-Minus and his number of possessions played.

___

Before analyzing player statistics we need to know how much each player is worth. Let's load individual player salaries. 

```{r, echo=FALSE}

# -----------------------------------------------------------------------------
# parse_season_table
#

parse_season_table <- function(table) {
  duplicated_header_rows <- table$RK == "RK"
  table <- table[!duplicated_header_rows, ]
  converted <- lapply(table, maybe_as_numeric)
  converted <- lapply(converted, empty_string_to_na)
  df <- as.data.frame(converted, stringsAsFactors = FALSE)
  
  # take position {PG, SG, SF, PF, C} from PlayerPos and make a new column called Pos
  df$POS <- str_split_fixed(df$NAME, ",", 2)[,2]
  
  # rename PlayerPos as Player
  df <- df %>%
    dplyr::rename(Player = NAME,
    )
  
  # Remove ", PG, SG, SF, PF, C}" from Player
  df$Player <- gsub("(.*),.*", "\\1", df$Player)
  
  
  return(df)
}



# -----------------------------------------------------------------------------
# get_espn_stats
#

get_espn_stats <- function() {
  for (i in 1:14) {
    if (i == 1) {
      url <- paste0("http://www.espn.com/nba/statistics/rpm")
    }
    else {
      url <- paste0("http://www.espn.com/nba/statistics/rpm/_/page/", i)
    }
    html <- xml2::read_html(url)
    node <- rvest::html_node(html, "table")
    table <- rvest::html_table(node, header = TRUE)
    if (i == 1) {
      df <- parse_season_table(table)
    }
    else {
      df <- rbind(df, parse_season_table(table))
    }
  }  
  return(df)
}
```


+ Web scraping hoopshype.com/salaries/players/...

```{r, echo=FALSE}

# ===========================================================================
# web scrape hoopshype.com for player salaries

#print("Web scraping hoopshype.com/salaries/players/...")
df_hoops_ps <- get_hoops_salary("https://hoopshype.com/salaries/players/")
ps_dim <- dim(df_hoops_ps)
paste0("Loaded ", ps_dim[1], " player salaries.")

```

#### Boxplot of NBA player salaries:

```{r, message=FALSE,warning=FALSE,echo=FALSE,fig.align='center', fig.width=3, fig.height=5}

ggplot(df_hoops_ps, aes(y=X2020.21/1000000)) +
  geom_boxplot(aes()) +
  theme_minimal() +
  theme(axis.text.x=element_blank()) +
  ylab("20-21 NBA Player Salaries ($M)")

ps_mean <- mean(df_hoops_ps$X2020.21)

```

<br><br><br><br>

##### Curious how much your favorite player is being paid this season? Search and see!

```{r, echo=FALSE, message=FALSE}
datatable(df_hoops_ps, rownames = FALSE, filter="top", options = list(pageLength = 5, scrollX=T) )
```

<br><br><br><br>

+ Web scraping espn.com/nba/statistics/...

```{r, echo=FALSE}
# ===========================================================================
# web scrape ESPN stats for GP, MPG, ORPM, DRPM, WINS
# Split "Name, Pos" into "Player" and "POS"

#print("Web scraping espn.com/nba/statistics/...")
df_espn <- get_espn_stats()

pe_dim <- dim(df_espn)
paste0("Loaded ", pe_dim[2], " statistics for ", pe_dim[1], " players.")

head(df_espn)
```

Add player salary as final column to player stats

```{r, echo=FALSE}
# Make new dataframe of stats and salaries
df_players <- dplyr::inner_join(df_espn, df_hoops_ps, by="Player")
#print("Add player salary as final column to player stats" )
head(df_players)
#str(df_players)
```

___

#### Analyzing RPM vs. Salary

```{r, message=FALSE,warning=FALSE,echo=FALSE,out.width=c('50%', '50%'), fig.show='hold'}

ggplot(df_players, aes(x=RPM, y=X2020.21/1000000)) +
  geom_point(aes(), show.legend=FALSE) +
  theme_minimal() +
  xlab("Real Plus-Minus") +
  ylab("Salary ($M)")

ggplot(df_players, aes(x=RPM, y=X2020.21/1000000, color=MPG, size=MPG)) +
  geom_point(aes(), alpha = 0.5, show.legend = TRUE) +
  theme_minimal() +
  scale_colour_gradientn(colors = rainbow(n=10)) +
  xlab("Real Plus-Minus") +
  ylab("Salary ($M)")
  
```

#### What's going on? 
The chart on the left is a plain scatterplot mess. On the right we can understand the data a little bit more. 

+ In general, <i>players that <b>play more</b> minutes are <b>paid more</b> and have a <b>higher plus/minus</b>.</i> This might make things difficult for us because we need a competitive team that is cheap!

___
## Time to collect more data


Our Plus/Minus stats come from [NBA.com](https://stats.nba.com/stats/leaguedashplayerstats).

Also, we will web scrape the data thanks to some help from our guy [Ashwin](https://towardsdatascience.com/how-scraping-nba-stats-is-cooler-than-michael-jordan-49d7562ce3ef). 

+ Web scraping stats.nba.com/stats/...

```{r, echo=FALSE, message=FALSE,warning=FALSE}



# -----------------------------------------------------------------------------
# get_nba_player_stats
#

get_nba_player_stats <- function() {

# https://stackoverflow.com/questions/46781563/how-to-obtain-a-json-response-from-the-stats-nba-com-api  
# https://towardsdatascience.com/how-scraping-nba-stats-is-cooler-than-michael-jordan-49d7562ce3ef  
  
  headers = c(
    `Connection` = 'keep-alive',
    `Accept` = 'application/json, text/plain, */*',
    `x-nba-stats-token` = 'true',
    `X-NewRelic-ID` = 'VQECWF5UChAHUlNTBwgBVw==',
    `User-Agent` = 'Mozilla/5.0 (Macintosh; Intel Mac OS X 11_2_3) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/78.0.3904.87 Safari/537.36',
    `x-nba-stats-origin` = 'stats',
    `Sec-Fetch-Site` = 'same-origin',
    `Sec-Fetch-Mode` = 'cors',
    `Referer` = 'https://stats.nba.com/players/shooting/',
    `Accept-Encoding` = 'gzip, deflate, br',
    `Accept-Language` = 'en-US,en;q=0.9'
  )
  
  url <- paste0("https://stats.nba.com/stats/leaguedashplayerstats",
                "?College=",
                "&Conference=",
                "&Country=",
                "&DateFrom=",
                "&DateTo=",
                "&Division=",
                "&DraftPick=",
                "&DraftYear=",
                "&GameScope=",
                "&GameSegment=",
                "&Height=",
                "&LastNGames=0",
                "&LeagueID=00",
                "&Location=",
                "&MeasureType=Base",
                "&Month=0",
                "&OpponentTeamID=0",
                "&Outcome=",
                "&PORound=0",
                "&PaceAdjust=N",
                "&PerMode=Totals",
                "&Period=0",
                "&PlayerExperience=",
                "&PlayerPosition=",
                "&PlusMinus=N",
                "&Rank=N",
                "&Season=2020-21",
                "&SeasonSegment=",
                "&SeasonType=Regular+Season",
                "&ShotClockRange=",
                "&StarterBench=",
                "&TeamID=0",
                "&TwoWay=0",
                "&VsConference=",
                "&VsDivision=",
                "&Weight=")
                
  res <- httr::GET(url = url, httr::add_headers(.headers=headers))
  json_resp <- jsonlite::fromJSON(httr::content(res, "text"))
  frame <- data.frame(json_resp$resultSets$rowSet)
  df <- data.frame(json_resp$resultSets$rowSet[1])
  colnames(df) <- json_resp[["resultSets"]][["headers"]][[1]]
  
  return(df)
}

```
```{r, echo=FALSE}
# ===========================================================================
# web scrape nba.com/stats for player general tradition stats

#print("Web scraping stats.nba.com/stats/...")
df_nba <- get_nba_player_stats()
pn_dim <- dim(df_nba)
paste0("Loaded ", pn_dim[2], " statistics for ", pn_dim[1], " players.")
#str(df_nba)
```


```{r, echo=FALSE, message=FALSE,warning=FALSE }
df_nba <- df_nba %>%
    dplyr::rename(Player = PLAYER_NAME,
    )
#str(df_nba)
```



```{r, echo=FALSE, message=FALSE,warning=FALSE}
# convert chr to integer
df_nba$AGE <- as.integer(df_nba$AGE)
df_nba$GP <- as.integer(df_nba$GP)
df_nba$W <- as.integer(df_nba$W)
df_nba$L <- as.integer(df_nba$L)
df_nba$FGM <- as.integer(df_nba$FGM)
df_nba$FGA <- as.integer(df_nba$FGA)
df_nba$FG3M <- as.integer(df_nba$FG3M)
df_nba$FG3A <- as.integer(df_nba$FG3A)
df_nba$FTM <- as.integer(df_nba$FTM)
df_nba$FTA <- as.integer(df_nba$FTA)
df_nba$OREB <- as.integer(df_nba$OREB)
df_nba$DREB <- as.integer(df_nba$DREB)
df_nba$REB <- as.integer(df_nba$REB)
df_nba$AST <- as.integer(df_nba$AST)
df_nba$TOV <- as.integer(df_nba$TOV)
df_nba$STL <- as.integer(df_nba$STL)
df_nba$BLK <- as.integer(df_nba$BLK)
df_nba$BLKA <- as.integer(df_nba$BLKA)
df_nba$PF <- as.integer(df_nba$PF)
df_nba$PFD <- as.integer(df_nba$PFD)
df_nba$PTS <- as.integer(df_nba$PTS)
df_nba$PLUS_MINUS <- as.integer(df_nba$PLUS_MINUS)
df_nba$DD2 <- as.integer(df_nba$DD2)
df_nba$TD3 <- as.integer(df_nba$TD3)


# convert chr to numeric
df_nba$W_PCT <- as.numeric(df_nba$W_PCT)
df_nba$FG_PCT <- as.numeric(df_nba$FG_PCT)
df_nba$FG3_PCT <- as.numeric(df_nba$FG3_PCT)
df_nba$FT_PCT <- as.numeric(df_nba$FT_PCT)
df_nba$NBA_FANTASY_PTS <- as.numeric(df_nba$NBA_FANTASY_PTS)


```


Now that we have all the individual player statistics we could ever need to analyze a player's value we will shift over to compile NBA team statistics. 

```{r, echo=FALSE, message=FALSE,warning=FALSE }

# Make new dataframe of stats and salaries

#dim(df_players)
df_p <- dplyr::inner_join(df_players, df_nba, by="Player")
#dim(df_p)
#head(df_p)

```

___


```{r, echo=FALSE }
# -----------------------------------------------------------------------------
# get_nba_records
#

get_nba_records <- function() {

  headers = c(
    `Connection` = 'keep-alive',
    `Accept` = 'application/json, text/plain, */*',
    `x-nba-stats-token` = 'true',
    `X-NewRelic-ID` = 'VQECWF5UChAHUlNTBwgBVw==',
    `User-Agent` = 'Mozilla/5.0 (Macintosh; Intel Mac OS X 11_2_3) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/78.0.3904.87 Safari/537.36',
    `x-nba-stats-origin` = 'stats',
    `Sec-Fetch-Site` = 'same-origin',
    `Sec-Fetch-Mode` = 'cors',
    `Referer` = 'https://www.nba.com/',
    `Accept-Encoding` = 'gzip, deflate, br',
    `Accept-Language` = 'en-US,en;q=0.9'
  )
  
  url <- paste0("https://stats.nba.com/stats/leaguestandingsv3",
                "?LeagueID=00",
                "&Season=2020-21",
                "&SeasonType=Regular%20Season")
                
  res <- httr::GET(url = url, httr::add_headers(.headers=headers))
  json_resp <- jsonlite::fromJSON(httr::content(res, "text"))
  frame <- data.frame(json_resp$resultSets$rowSet)
  df <- data.frame(json_resp$resultSets$rowSet[1])
  colnames(df) <- json_resp[["resultSets"]][["headers"]][[1]]
  
  return(df)
}

```


+ Web scraping nba.com/standings...


```{r, echo=FALSE}
# ===========================================================================
# web scrape ESPN stats for GP, MPG, ORPM, DRPM, WINS
# Split "Name, Pos" into "Player" and "POS"

#print("Web scraping nba.com/standings ...")
df_tr <- get_nba_records()

#str(df_tr)

tr_dim <- dim(df_tr)
paste0("Loaded ",tr_dim[2], " stats for ", tr_dim[1], " teams.")


# convert chr to integer
df_tr$WINS <- as.integer(df_tr$WINS)
df_tr$LOSSES <- as.integer(df_tr$LOSSES)

# convert chr to numeric
df_tr$WinPCT <- as.numeric(df_tr$WinPCT)
df_tr$PointsPG <- as.numeric(df_tr$PointsPG)
df_tr$OppPointsPG <- as.numeric(df_tr$OppPointsPG)
df_tr$DiffPointsPG <- as.numeric(df_tr$DiffPointsPG)


# rename TeamCity as Team
# rename WINS as W, LOSSES as L
# rename WinPCT as W_PCT,
# rename PointsPG = PPG
# rename OppPointsPG = OppPPG
df_tr <- df_tr %>% 
  dplyr::rename(Team = TeamCity,
                W = WINS,
                L = LOSSES,
                W_Pct = WinPCT,
                PPG = PointsPG,
                OppPPG = OppPointsPG)

#dim(df_tr)

# select only the columns we need: Team, Team Name, 
#                                  W, L, and 
#                                  PPG, OPPG, DiffPPG
df_tr <- df_tr %>% dplyr::select(Team, TeamName, 
                                 W, L, 
                                 PPG:DiffPointsPG) 

#dim(df_tr)

# Note: NBA.com uses the following convention:
# Team        | Name
# ------------+--------------
# LA          | Clippers
# Los Angeles | Lakers
#
# hoopshype.com uses the following convention:
# Team        
# -------------
# LA Clippers
# LA Lakers


# Resolve name issue for 2 LA teams
df_hoops_ts$Team[df_hoops_ts$Team == "LA Clippers"] <- "LA"
df_hoops_ts$Team[df_hoops_ts$Team == "LA Lakers"] <- "Los Angeles"

#dim(df_tr)

df_tr <- dplyr::inner_join(df_tr, df_hoops_ts, by="Team")
#dim(df_tr)

df_t <- df_tr
t_dim <- dim(df_t)
paste0("Cleaned and reduced to ",t_dim[2], 
       " stats (including salary) for ", t_dim[1], " teams.")

#head(df_t)
```

##### Why the team stats? 
These stats will be enable us to test how competitively our team stacks up against other teams in the NBA. 

```{r, echo=FALSE, message=FALSE}
head(df_t)
```

___

#### Team Salary vs. Team Wins
Let's see if spending more money translates to winning more games.   

```{r, message=FALSE,warning=FALSE,echo=FALSE,fig.align='center', fig.width=8, fig.height=6}

ggplot(df_t, aes(x=W, y=X2020.21/1000000, 
                 color = DiffPointsPG, size = DiffPointsPG)) +
  geom_point(aes(), show.legend = TRUE) +
  theme_minimal() +
  scale_color_gradient(low = "red", high = "green") +

  xlab("Team Wins") +
  ylab("Team Salary ($M)")

  
```

<br><br><br><br>

Looking at the data another way we observe a <b>pattern</b> that might be obvious. 
<br><br><br><br>

#### Team Wins vs. Team Diff PPG

```{r, message=FALSE,warning=FALSE,echo=FALSE,fig.align='center', fig.width=8, fig.height=6}

ggplot(df_t, aes(x=DiffPointsPG, y=W, 
                 color = X2020.21/1000000, size = X2020.21/1000000)) +
  geom_point(aes(), alpha = 1, show.legend = FALSE) +
  geom_smooth(method='lm') +
  theme_minimal() +
  scale_color_gradient(low = "green", high = "red") +
  xlab("Team DiffPointPG") +
  ylab("Wins") +
  guides(color=guide_legend(title="Salary ($M)")) +
  guides(size=guide_legend(title="Salary ($M)"))


  
```


### Wins and DiffPointPG
+ These stats are generally correlated
+ Higher DiffPointPG translates to more Wins
+ However, higher salary doesn't always correlate with more wins
  + There are lower salary teams with better PPG and Wins!

This finding gives us hope. It opens the door a little bit more for being able to <i>build a winning team</i> on a low budget. 

___

### Now what?
Well, we have evidence that it is possible to be a competitive team with a lower end budget. Let's see how far we can stretch the limits now that we have all the player and team statistics we need for analysis. 

___

#### First, let's filter out players...
+ With few minutes (we need players!)
+ And a Plus/Minus < 200 (we need good players)
+ With salaries over $15M (we need affordable players)



```{r, message=FALSE,warning=FALSE,echo=FALSE,fig.align='center', fig.width=8, fig.height=6}

# find highest Plus-Minus players that are affordable

#glimpse(df_p)

df_p2 <- df_p %>% 
  filter(MPG > 15) %>%
  filter(PLUS_MINUS > 200) %>%
  filter(X2020.21 < 15000000) %>%
  mutate(P_M_PG = PLUS_MINUS / GP.x)

paste0("Now we have ", nrow(df_p2)," players to analyze.")

ggplot(df_p2, aes(x=PLUS_MINUS, y=X2020.21/1000000,
                  color=MPG, size=MPG)) +
  geom_point(aes(), show.legend = TRUE) +
  theme_minimal() +
  scale_color_gradient(low = "yellow", high = "green") +
  xlab("Plus/Minus") +
  ylab("Salary ($M)")



roster <- df_p2 %>%
  select(Player, POS, MPG, P_M_PG, X2020.21) %>%
  arrange(X2020.21)

roster <- roster[1:12, ]

paste0("Here is the composition of the team:")
table(roster$POS)

paste0("Here is the team:")
roster



paste0("The salary is ", round(sum(roster$X2020.21/1000000),1), " M")

```

From the analysis of plus/minus we filtered out a team that is slightly outside of our budget and guard heavy. Not bad for a first attempt. Let's keep going with real plus minus. 

___

#### Attempt 2: let's filter out players...
+ With few minutes (we need players!)
+ And a Real Plus Minus < 1.8 (we need good players)
+ With salaries over $15M (we need affordable players)

```{r, message=FALSE,warning=FALSE,echo=FALSE,fig.align='center', fig.width=8, fig.height=6}

# find highest RPM players that are affordable

df_p3 <- df_p %>% 
  filter(MPG > 15) %>%
  filter(RPM > 1.8) %>%
  filter(X2020.21 < 15000000)

paste0("Now we have ", nrow(df_p3)," players to analyze.")

ggplot(df_p3, aes(x=RPM, y=X2020.21/1000000,
                  color=MPG, size=MPG)) +
  geom_point(aes(), show.legend = TRUE) +
  theme_minimal() +
  scale_color_gradient(low = "yellow", high = "green") +
  xlab("Real Plus Minus") +
  ylab("Salary ($M)")

df_p3table <- df_p3 %>%
  select(Player, POS, MPG, RPM, X2020.21) %>%
  arrange(X2020.21)

df_p3table <- df_p3table[1:12, ]

paste0("Here is the composition of the team:")
table(df_p3table$POS)

paste0("Here is the team:")
df_p3table

paste0("The salary is ", round(sum(df_p3table$X2020.21/1000000),1), " M")


```

From the analysis of real plus/minus, again, we filtered out a team that is slightly outside of our budget, but looking more like a normal lineup. Not bad for a second attempt. Let's keep going with real plus/minus wins. 


___

#### Attempt 3: let's filter out players...
+ With few minutes (we need players!)
+ And a RPM WINS < 5 (we need good players)
+ With salaries over $15M (we need affordable players)


```{r, message=FALSE,warning=FALSE,echo=FALSE,fig.align='center', fig.width=8, fig.height=6}

# find highest WINS players that are affordable

df_p4 <- df_p %>% 
  filter(MPG > 15) %>%
  filter(WINS > 5) %>%
  filter(X2020.21 < 15000000) %>%
  mutate(P_M_PG = PLUS_MINUS / GP.x)

paste0("Now we have ", nrow(df_p4)," players to analyze.")

ggplot(df_p4, aes(x=WINS, y=X2020.21/1000000,
                  color=MPG, size=MPG)) +
  geom_point(aes(), show.legend = TRUE) +
  theme_minimal() +
  scale_color_gradient(low = "yellow", high = "green") +
  xlab("WINS") +
  ylab("Salary ($M)")

df_p4table <- df_p4 %>%
  select(Player, POS, MPG, WINS, X2020.21) %>%
  arrange(X2020.21)

df_p4table <- df_p4table[1:12, ]

paste0("Here is the composition of the team:")
table(df_p4table$POS)

paste0("Here is the team:")
df_p4table

paste0("The salary is ", round(sum(df_p4table$X2020.21/1000000),1), " M")


```

#### Ayyoo! 

Now we are in business. Out of the 24 players selected we chose to stick with the 12 players that were the least expensive. This landed us right where we projected we could be with a competitive team.  


___

## Now, the real question: 
### Can this newly created $Bball team compete with current teams?
+ Let's test by pro-rating RPM WINS vs. existing league leaders' RPM WINS

```{r, message=FALSE,warning=FALSE,echo=FALSE,fig.align='center', fig.width=8, fig.height=6}


df_t <- df_t %>% 
  arrange(desc(W))

paste0("The top  team is: ", df_t$Team[1], " with ", df_t$W[1], " wins and ", df_t$L[1], " losses.")

#glimpse(df_p)
#glimpse(df_t)

df_t1 <- df_p %>%
  filter(TEAM == "PHX") %>%
  select(TEAM, Player, GP.x, MPG, PLUS_MINUS, RPM, WINS)

#df_t1

# WINS: "RPM Wins" provide an estimate of the number of wins each player has
#       contributed to his team's win total on the season. RPM Wins include the
#       player's Real Plus-Minus and his number of possessions played.

t1_sum_wins <- round(sum(df_t1$WINS))

#paste0(df_t$Team[1], " roster adds up to ", t1_sum_wins, " wins.")

games_played <- df_t$W[1] + df_t$L[1]

# games_played = 64 as of 5/3 6pm Mountain

df_p4table <- df_p4table %>%
  arrange(desc(WINS))

# pro-rate MPG and hence WINS based on projected minutes
# use highest WINS players as follows:
# 1-5 36 min
# 6-10 12 min
# 11-12 0 min

df_p4munge <- df_p4table

df_p4munge <- df_p4munge %>%
  mutate(pr_MPG = 34)

df_p4munge$pr_MPG[4:8] <- 24

df_p4munge$pr_MPG[9:10] <- 9

df_p4munge$pr_MPG[11:12] <- 0

#df_p4munge

df_p4munge <- df_p4munge %>%
  mutate(pr_WINS = WINS * pr_MPG / MPG)

#df_p4munge

p4_sum_wins <- round(sum(df_p4munge$pr_WINS))
  
```

___

#### Our team will play a similar style to other NBA teams:
+ players 1-3 will play 34 minutes per game (mpg),
+ players 4-8 will play 24 mpg,
+ players 9 & 10 will play 9 mpg, 
+ and players 11 & 12 will be reserves. 
```{r, message=FALSE,warning=FALSE,echo=FALSE,fig.align='center', fig.width=8, fig.height=6}

df_t_new <- df_t %>% 
  add_row(Team = "SmallMarket", 
          TeamName = "$Bball",
          W = p4_sum_wins,
          L = games_played - p4_sum_wins,  
          PPG = NA,
          OppPPG = NA,
          DiffPointsPG = NA,
          X2020.21 = sum(df_p4table$X2020.21))

ggplot(df_t_new, aes(x=W, y=X2020.21/1000000, color = W, size = 5)) +
  geom_point(aes(), show.legend = FALSE) +
  theme_minimal() +
  scale_color_gradient(low = "red", high = "green") +
  xlab("Team Wins") +
  ylab("Team Salary ($M)")

paste0("Our $Bball team, with a salary of ",
       round(sum(df_p4table$X2020.21/1000000),1),
       "M, has a projected W-L record of ", 
       p4_sum_wins, "-", games_played - p4_sum_wins, 
       "!!!")
  
```

___

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

## Success! 
###### We are on track to head into the playoffs as the <b>number 1 overall team</b>! At this rate we may even post the best record of all-time. 

</div>

___











