---
title: "UVU Offense"
date: "May 10, 2021"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
pagetitle: Austin Daines
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(modelr)
library(plyr)
library(readxl)
library(dplyr)
df1 <- read_excel(path = "OFFENSE.xlsx", sheet = "CBU")
df2 <- read_excel(path = "OFFENSE.xlsx", sheet = "CBU2")
df3 <- read_excel(path = "OFFENSE.xlsx", sheet = "Seattle")
df4 <- read_excel(path = "OFFENSE.xlsx", sheet = "Tarleton")
df5 <- read_excel(path = "OFFENSE.xlsx", sheet = "Tarleton 2")
df6 <- read_excel(path = "OFFENSE.xlsx", sheet = "Dixie St.")
df7 <- read_excel(path = "OFFENSE.xlsx", sheet = "Dixie St. 2")
df8 <- read_excel(path = "OFFENSE.xlsx", sheet = "NMSU")
df9 <- read_excel(path = "OFFENSE.xlsx", sheet = "NMSU 2")
df10 <- read_excel(path = "OFFENSE.xlsx", sheet = "UTRGV")
df11 <- read_excel(path = "OFFENSE.xlsx", sheet = "UTRGV 2")
df12 <- read_excel(path = "OFFENSE.xlsx", sheet = "GCU")
df13 <- read_excel(path = "OFFENSE.xlsx", sheet = "GCU 2")
df14 <- read_excel(path = "OFFENSE.xlsx", sheet = "NMSU3")
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
df1$Opponent <- "CBU"
df2$Opponent <- "CBU2"
df3$Opponent <- "SU"
df4$Opponent <- "TSU"
df5$Opponent <- "TSU2"
df6$Opponent <- "DSU"
df7$Opponent <- "DSU2"
df8$Opponent <- "NMSU"
df9$Opponent <- "NMSU2"
df10$Opponent <- "UTRGV"
df11$Opponent <- "UTRGV2"
df12$Opponent <- "GCU"
df13$Opponent <- "GCU2"
df14$Opponent <- "NMSU3"
df2$...8 <- NULL
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
conf <- rbind(df1,df2,df3,df4,df5,df6,df7,df8,df9, df10, df11, df12, df13, df14)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
names(conf)[7] <- "pts"
names(conf)[2] <- "passes"
names(conf)[3] <- "sides"

unique(conf$pts)

conf$pts[conf$pts == "TO"] <- "0"
conf$pts[conf$pts == "Miss2pt"] <- "0"
conf$pts[conf$pts == "Miss3pt"] <- "0"
conf$pts[conf$pts == "Make2pt"] <- "2"
conf$pts[conf$pts == "Make3pt"] <- "3"
conf$pts[conf$pts == "miss2"] <- "0"
conf$pts[conf$pts == "miss3"] <- "0"
conf$pts[conf$pts == "make2"] <- "2"
conf$pts[conf$pts == "make3"] <- "3"

conf$pts[conf$pts == "miss2pt"] <- "0"
conf$pts[conf$pts == "miss3pt"] <- "0"
conf$pts[conf$pts == "make2pt"] <- "2"
conf$pts[conf$pts == "miss 2pt"] <- "0"
conf$pts[conf$pts == "miss23pt"] <- "0"
conf$pts[conf$pts == "made2"] <- "2"
conf$pts[conf$pts == "made3"] <- "3"

conf$pts[conf$pts == "Miss23pt"] <- "0"
conf$pts[conf$pts == "Miss 2pt"] <- "0"
conf$pts[conf$pts == "FOUL"] <- "1"
conf$pts[conf$pts == "Shooting foul"] <- "1"
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
new_conf <- conf
unique(conf$pts)
class(conf$pts)
class(conf$passes)
class(conf$sides)

conf$pts = as.numeric(conf$pts)
conf$passes = as.numeric(conf$passes)
conf$sides = as.numeric(conf$sides)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
pass <- table(new_conf$passes)
pass <- as.data.frame(pass)
names(pass)[1] <- 'num_of_passes'
str(new_conf)
new_conf$passes <- as.numeric(new_conf$passes)

```


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#points per number of passes
half_court <- conf[!(conf$play=="Transition"),]
half_court <- half_court[ , c("play","passes","sides","pts")]

half_court %>% group_by(passes, sides) %>% 
    summarize(total_points = sum(pts, na.rm = TRUE))

table(half_court$passes)

```


# Ball Movement and Scoring 

<span style="color:black"> During the 2020-21 season the <span style="color:green"> **Utah Valley Wolverines** </span> ran a motion offense that regularly moved the ball from one side of the court to the other and back. When timing was right and ball screen, cuts, and specific actions were executed the offense worked well. </span>

<span style="color:black"> This motion offense begged the question: *Was all the side to side movement and passing contributing to offensive success?* </span>

<span style="color:black"> A quick analysis might help us find out. </span>

<br>

# The Gathering 

<span style="color:black"> Conference play started in January 2021 and I was asked to record each offensive possession and tally </span>

  * <span style="color:black"> the number of passes once was the ball was in the half court. </span>
  * <span style="color:black"> the number of times the ball switched sides of the half court. </span>

<span style="color:black"> These two stats are not tracked on ESPN or any college basketball platform that I am aware of so I would re-watch each game and manually track these occurrences. </span>

<br>

# The Visuals


```{r, echo=FALSE, message=FALSE, warning=FALSE, dpi = 200}
pass_bar <- ggplot(half_court, aes(x=passes)) +
  geom_bar(color="black", fill="dark green") +
  theme_minimal() +
  xlab("\nNumber of Passes") +
  ylab("Number of Possessions\n") +
  #ggtitle("Passes on Offensive Possessions\n",
          #subtitle = "Conference Games 2020-21 Season") +
  theme(plot.title = element_text(size = 26, face = "bold"),
        plot.subtitle = element_text(vjust = 14.0)) +
  theme(axis.text=element_text(size=12), 
        axis.title = element_text(size=14))
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, dpi = 200}
sides_bar <- ggplot(half_court, aes(x=sides)) +
  geom_bar(color="black", fill="dark green") +
  theme_minimal() +
  xlab("\nNumber of Sides") +
  ylab("Number of Possessions\n") +
  #ggtitle("Sides on Offensive Possessions\n",
          #subtitle = "Conference Games 2020-21 Season") +
  theme(plot.title = element_text(size = 26, face = "bold"),
        plot.subtitle = element_text(vjust = 14.0)) +
  theme(axis.text=element_text(size=12), 
        axis.title = element_text(size=14))
```

### Passing

```{r, echo=FALSE, message=FALSE, warning=FALSE, dpi = 200}
pass_bar <- pass_bar + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

pass_bar
```

<br>

### Ball Movement

```{r, echo=FALSE, message=FALSE, warning=FALSE, dpi = 200}
sides_bar <- sides_bar + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

sides_bar
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#points per number of passes
str(half_court)

half_court[order(half_court$passes, half_court$sides, half_court$pts, decreasing = TRUE),]

?order()
table(half_court$passes)
table(half_court$sides)

half_court <- half_court[complete.cases(half_court), ]

sum(half_court[which(half_court$passes=='0'), 4])
sum(half_court[which(half_court$passes=='1'), 4])
sum(half_court[which(half_court$passes=='2'), 4])
sum(half_court[which(half_court$passes=='3'), 4])
sum(half_court[which(half_court$passes=='4'), 4])
sum(half_court[which(half_court$passes=='5'), 4])
sum(half_court[which(half_court$passes=='6'), 4])
sum(half_court[which(half_court$passes=='7'), 4])
sum(half_court[which(half_court$passes=='8'), 4])
sum(half_court[which(half_court$passes=='9'), 4])
sum(half_court[which(half_court$passes=='10'), 4])
sum(half_court[which(half_court$passes=='11'), 4])

poss <- table(half_court$passes)
# 0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11
# 60, 158, 56, 122, 61, 62, 44, 32, 5, 8, 3, 0

total_pts <- data.frame(passes = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10), 
                        points = c(60, 158, 56, 122, 61, 62, 44, 32, 5, 8, 3))
poss <- as.data.frame(poss)
poss <- poss[0:11,]
ppp <-  round(total_pts$points / poss$Freq, digits = 2)

ppp <- data.frame(passes = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
                  ppp)
str(ppp)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, dpi = 200}
pass_ppp <- ggplot(ppp, aes(x=passes, y=ppp)) +
  geom_point(color="dark green", size = 2.5) +
  geom_smooth(method = "lm", fill=NA, color="black") +
  theme_minimal() +
  xlab("\nNumber of Passes") +
  ylab("Points Per Possessions\n") +
  #ggtitle("PPP vs. Passes\n",
            #subtitle = "Conference Games 2020-21 Season") +
  theme(plot.title = element_text(size = 26, face = "bold"),
        plot.subtitle = element_text(vjust = 14.0)) +
  theme(axis.text=element_text(size=12), 
        axis.title = element_text(size=14))
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#points per number of sides
sum(half_court[which(half_court$sides=='0'), 4])
sum(half_court[which(half_court$sides=='1'), 4])
sum(half_court[which(half_court$sides=='2'), 4])
sum(half_court[which(half_court$sides=='3'), 4])
sum(half_court[which(half_court$sides=='4'), 4])
sum(half_court[which(half_court$sides=='5'), 4])
sum(half_court[which(half_court$sides=='6'), 4])
sum(half_court[which(half_court$sides=='7'), 4])

poss_sides <- table(half_court$sides)

total_pts_sides <- data.frame(sides = c(1, 2, 3, 4, 5), 
                        points = c(153, 202, 141, 76, 32))
poss_sides <- as.data.frame(poss_sides)
poss_sides <- poss_sides[2:6,]
ppp_sides <-  round(total_pts_sides$points / poss_sides$Freq, digits = 2)

ppp_sides <- data.frame(sides = c(1, 2, 3, 4, 5),
                  ppp_sides)

```


```{r, echo=FALSE, message=FALSE, warning=FALSE, dpi = 200}
sides_ppp <- ggplot(ppp_sides, aes(x=sides, y=ppp_sides)) +
  geom_point(color="dark green", size = 2.5) +
  geom_smooth(method = "lm", fill=NA, color="black") +
  theme_minimal() + 
  xlab("\nNumber of Sides") +
  ylab("Points Per Possessions\n") #+
  #ggtitle("PPP vs. Sides\n",
          #subtitle = "Conference Games 2020-21 Season") +
  #theme(plot.title = element_text(size = 26, face = "bold"),
        #plot.subtitle = element_text(vjust = 14.0)) +
  #theme(axis.text=element_text(size=12), 
        #axis.title = element_text(size=14))
```


# The Scoring

<br>

### Passing vs. PPP

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE, dpi = 200}
pass_ppp <- pass_ppp + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

pass_ppp <- pass_ppp + scale_x_discrete(limits=c("1","2", "3", "4", "5", "6", "7", "8", "9", "10"))

pass_ppp
```


<br>


### Ball Movement vs. PPP


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE, dpi = 200}
sides_ppp <- sides_ppp + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

sides_ppp
```

<br>

# Quick Takeaways

* <span style="color:black"> On average an odd number of passes yielded more points per possession than an even number of passes. </span>

* <span style="color:black"> The number of sides displayed an nice bell-like curve in points per possession. </span>


```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#points per play

```

