---
title: "Summer Shooting -- 2021"
date: "April 2, 2022"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
pagetitle: Austin Daines
---

```{r, include=FALSE}
# load libraries
library(emojifont)
library(googlesheets4)
library(tidyverse)
```


```{r, include=FALSE}
# https://bigskyconf.com/sports/mbball
# grab df from google sheet
gs4_deauth()
link22 <- "https://docs.google.com/spreadsheets/d/1YIMkvigzO-ermQiLu4DOTsW9U7ThaW4fXV9Dc0uRYZs/edit#gid=0"
df22 <- read_sheet(link22)
```

```{r, include=FALSE}
# extract columns for each season
df21 <- df22 %>% select(5:7)
df22 <- df22 %>% select(2:4)
```

```{r, include=FALSE}
df21$three_fga[df21$three_fga == -9] <- 9

df21 <- df21 %>% mutate(running_pct = round(cumsum(three_fg) / cumsum(three_fga), 3) * 100)
df21$season <- "2020-21"

df22 <- df22 %>% mutate(running_pct = round(cumsum(three_makes) / cumsum(three_attempts), 3) * 100)
df22$season <- "2021-22"
```

# Three Point Shooting `r emoji("basketball")`

<span style="color:black"> In order to stay relevant at any level of basketball teams need to be able to shoot the ball well, especially from beyond the arc. </span> 

<span style="color:orange"> **Idaho State** </span> <span style="color:black"> ranked 9th out of 11 teams in the Big Sky conference for 3pt shooting percentage during the 2020-21 season at 33.3% on almost 21 attempts per game. </span>

<span style="color:black"> Three point shooting is not the only thing that helps a team win, but it has become a relevant factor. The top 3 finishers in the Big Sky conference standings ranked 2nd, 3rd, and 4th, in 3pt FG makes per game in 2020-21 season. Now, I'm not suggesting their shooting prowess was *the* reason they had great success during conference play, but it is intriguing relationship. </span>


<br>

# Summer Shooting

<span style="color:black"> I was tasked with tracking and recording (and later visualizing and analyzing) tens of thousands of three point shots over the course of the summer. <span style="color:orange"> **ISU's** </span> players devoted dozens of hours in the gym with one goal: **get better at shooting the 3 ball**. </span> 

* <span style="color:red"> **Spoiler Alert:** </span> <span style="color:black"> Although most players saw an [increase in performance](https://daisydaines.github.io/isu_mbb/summer_charts/) over the course of the summer, the team as whole did not improve three point shooting during the subsequent 2021-2022 season compared to the previous 2020-2021 season `r emoji("disappointed")`</span>

<br>

