---
title: "Summer Shooting -- 2021"
date: "April 2, 2022"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
pagetitle: Austin Daines
---

```{r, include=FALSE}
# load libraries
library(emojifont)
library(googlesheets4)
library(tidyverse)
```


```{r, include=FALSE}
# https://bigskyconf.com/sports/mbball
# grab df from google sheet
# deauth --anyone with link can access
gs4_deauth()

# url
link22 <- "https://docs.google.com/spreadsheets/d/1YIMkvigzO-ermQiLu4DOTsW9U7ThaW4fXV9Dc0uRYZs/edit#gid=0"
df <- read_sheet(link22)
```

```{r, include=FALSE}
# extract columns for each season
# 2019-20
df20 <- df %>% select(8:10)
# 2020-21
df21 <- df %>% select(5:7)
# 2021-22
df22 <- df %>% select(2:4)
```


```{r, include=FALSE}
# rename columns
df20 <- df20 %>% rename(
  three_fg = three_fg...8,
  three_fga = three_fga...9,
  three_pct = three_pct...10
)

df21 <- df21 %>% rename(
  three_fg = three_fg...5,
  three_fga = three_fga...6,
  three_pct = three_pct...7
)

# print column names
colnames(df20)
colnames(df21)
```


```{r, include=FALSE}
# check structure
str(df20)
str(df21)
str(df22)
# change columns to numeric 
df20$three_fga <- as.numeric(df20$three_fga)
df21$three_fga <- as.numeric(df21$three_fga)
df22$three_attempts <- as.numeric(df22$three_attempts)
# add column for games
df20$game <- 1:nrow(df20)
df21$game <- 1:nrow(df21)
df22$game <- 1:nrow(df22)

# add running total for percentage
df20 <- df20 %>% mutate(running_pct = round(cumsum(three_fg) / cumsum(three_fga), 3) * 100)
#df21 <- df21 %>% mutate(running_pct = round(cumsum(three_fg) / cumsum(three_fga), 3) * 100)
df22 <- df22 %>% mutate(running_pct = round(cumsum(three_makes) / cumsum(three_attempts), 3) * 100)
```


```{r, include=FALSE}
# remove last 6 rows NA data
df21 <- head(df21, - 6)
# fix two numbers
df21[df21$three_fga== -9, "three_fga"] <- 9
df20[df20$three_fga == -7, "three_fga"] <- 7
```


# Three Point Shooting `r emoji("basketball")`

<span style="color:black"> The three point shot has taken basketball, at every level, by storm over the last decade or so. It seems that the teams that are able to compete are the ones team cashing in on this valuable shot beyond the arc.  </span> 

<span style="color:orange"> **Idaho State** </span> <span style="color:black"> ranked 9th out of 11 teams in the Big Sky conference for 3pt shooting percentage during the 2020-21 season at 33.3% on almost 21 attempts per game. Also, the <span style="color:orange"> **Bengals** </span> ranked 8th out of 11 teams in 3 point field goals made per game. </span>

<span style="color:black"> Three point shooting is not the only thing that helps a team win, but it has become an important factor. The top 3 finishers in the Big Sky conference standings ranked 2nd, 3rd, and 4th, in 3pt FG makes per game in 2020-21 season. Now, I'm not suggesting their shooting prowess was *the* reason they had great success during conference play, but it is intriguing relationship. </span>


<br>

# Summer Shooting

<span style="color:black"> I was tasked with tracking and recording (and later visualizing and analyzing) tens of thousands of three point shots over the course of the summer. <span style="color:orange"> **ISU's** </span> players devoted dozens of hours in the gym with one goal: **get better at shooting the 3 ball**. </span> 

* <span style="color:red"> **Spoiler Alert:** </span> <span style="color:black"> Although all players saw an [increase in performance](https://daisydaines.github.io/isu_mbb/summer_charts/) over the course of the summer, the team did not improve overall three point shooting during the subsequent 2021-22 season compared with the previous two seasons. `r emoji("disappointed")`</span>

<br>


```{r, include=FALSE}
# build simple linear model to predict final six games for COVID shortened season
model <- lm(three_fg ~ three_fga, df21)
# build df
pred_df <- data.frame(three_fga = floor(runif(6, min= min(df21$three_fga), max= max(df21$three_fga))), 
                         game = 25:30)
# add predicted numbers
pred_df$three_fg <- predict(model, pred_df)
```



```{r, include=FALSE}
# add original columns to later bind
str(pred_df)
pred_df$three_pct <- pred_df$three_fg / pred_df$three_fga
```


```{r, include=FALSE}
# reorder columns to match for rbind
colnames(pred_df)
pred_df <- pred_df[,c(3,1,4,2)]
```


```{r, include=FALSE}
# rbind predictions with df21
df21 <- rbind(df21, pred_df)
```


```{r, include=FALSE}
# add running percentage for entire season 
df21 <- df21 %>% mutate(running_pct = round(cumsum(three_fg) / cumsum(three_fga), 3) * 100)
```


```{r, include=FALSE}
# rename and reorder df22 columns
colnames(df22)
df22 <- df22[, c(1,2,3,5,4)]

df22 <- df22 %>% rename(
  three_fg = three_makes,
  three_fga = three_attempts,
  three_pct = pct
)
```


```{r, include=FALSE}
# add column for season to group by in ggplot
df20$season <- "2019-20"
df21$season <- "2020-21"
df22$season <- "2021-22"
# bind all dfs for dfall
dfall <- rbind(df20, df21, df22)
```



```{r, include=FALSE}
# ggplot: line chart colored by season
three <- ggplot(dfall, aes(x=game, y=running_pct, group=season), size = 1.5) +
  geom_line(aes(color=season), size= 1) + 
  theme_minimal() + 
  labs(y = "Running 3pt Percentage", 
       x= "Game")
# add relevant ISU colors
three <- three + scale_color_manual(breaks = c("2019-20", "2020-21", "2021-22"),
                        values=c("gray", "orange", "black"))
```


# Season Comparison

```{r, echo=FALSE, dpi = 200}
# remove grid lines and make outer lines black for simple looks
three <- three + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
# print plot
print(three)
```


# Quick Thoughts

* <span style="color:black"> The difference between 33% and 31% may seem trivial, but 2% on almost 700 3pt attempts during the season adds up quickly. </span>

* <span style="color:black"> Finishing above 33% for the 2021-22 season would have meant making 15+ more threes. That's an extra three pointer every other game. </span>

* <span style="color:orange"> **Idaho State** </span> <span style="color:black"> finished the season 7-23. </span>
    * <span style="color:black"> 10 of the 23 losses came by fewer than 9 points </span>
    * <span style="color:black"> 6 of those 10 single digit losses were by 5 points or fewer.  </span>

* <span style="color:black"> An extra three here or there might have been the difference between a dismal 7-23 season or an ok 13-17 season `r emoji("pensive")` </span>

