---
title: "Summer Shooting -- 2021"
date: "April 2, 2022"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
pagetitle: Austin Daines
---

```{r, include=FALSE}
# load libraries
library(emojifont)
library(googlesheets4)
library(tidyverse)
```


```{r, include=FALSE}
# https://bigskyconf.com/sports/mbball
# grab df from google sheet
gs4_deauth()
link22 <- "https://docs.google.com/spreadsheets/d/1YIMkvigzO-ermQiLu4DOTsW9U7ThaW4fXV9Dc0uRYZs/edit#gid=0"
df22 <- read_sheet(link22)
```

```{r, include=FALSE}
# extract columns for each season
df21 <- df22 %>% select(5:7)
df22 <- df22 %>% select(2:4)
```

```{r, include=FALSE}
df21$three_fga[df21$three_fga == -9] <- 9

#df21 <- df21 %>% mutate(running_pct = round(cumsum(three_fg) / cumsum(three_fga), 3) * 100)
df21$game <- 1:nrow(df21)

#df21 <- head(df21, - 6)
df21$three_fga <- as.numeric(df21$three_fga)

df22 <- df22 %>% mutate(running_pct = round(cumsum(three_makes) / cumsum(three_attempts), 3) * 100)
df22$game <- 1:nrow(df22)

# convert three_fga to num
df21 <- head(df21, - 6)
df21$three_fga <- as.numeric(df21$three_fga)
```

# Three Point Shooting `r emoji("basketball")`

<span style="color:black"> In order to stay relevant at any level of basketball teams need to be able to shoot the ball well, especially from beyond the arc. </span> 

<span style="color:orange"> **Idaho State** </span> <span style="color:black"> ranked 9th out of 11 teams in the Big Sky conference for 3pt shooting percentage during the 2020-21 season at 33.3% on almost 21 attempts per game. </span>

<span style="color:black"> Three point shooting is not the only thing that helps a team win, but it has become a relevant factor. The top 3 finishers in the Big Sky conference standings ranked 2nd, 3rd, and 4th, in 3pt FG makes per game in 2020-21 season. Now, I'm not suggesting their shooting prowess was *the* reason they had great success during conference play, but it is intriguing relationship. </span>


<br>

# Summer Shooting

<span style="color:black"> I was tasked with tracking and recording (and later visualizing and analyzing) tens of thousands of three point shots over the course of the summer. <span style="color:orange"> **ISU's** </span> players devoted dozens of hours in the gym with one goal: **get better at shooting the 3 ball**. </span> 

* <span style="color:red"> **Spoiler Alert:** </span> <span style="color:black"> Although most players saw an [increase in performance](https://daisydaines.github.io/isu_mbb/summer_charts/) over the course of the summer, the team as whole did not improve three point shooting during the subsequent 2021-22 season compared with the previous season `r emoji("disappointed")`</span>

<br>


```{r, include=FALSE}
#model <- lm(running_pct ~ game, df21)
#predictDf <- data.frame(game = 25:30)
#predictDf$running_pct <- predict(model, predictDf)

model <- lm(three_fg ~ three_fga, df21)
pred_df <- data.frame(#three_fg = floor(runif(6, min= min(df21$three_fg), max= max(df21$three_fg))), 
                         three_fga = floor(runif(6, min= min(df21$three_fga), max= max(df21$three_fga))), 
                         #three_pct = three_fg / three_fga, 
                         #running_pct = round(cumsum(three_fg) / cumsum(three_fga), 3) * 100, 
                         game = 25:30)
pred_df$three_fg <- predict(model, pred_df)


```



```{r, include=FALSE}
# add original columns to later bind
str(pred_df)
pred_df$three_pct <- pred_df$three_fg / pred_df$three_fga
#pred_df$running_pct <- round(cumsum(pred_df$three_fg) / cumsum(pred_df$three_fga), 3) * 100
```


```{r, include=FALSE}
colnames(pred_df)
pred_df <- pred_df[,c(3,1,4,2)]
```

```{r, include=FALSE}
# rbind
df21 <- rbind(df21, pred_df)
```


```{r, include=FALSE}
df21 <- df21 %>% mutate(running_pct = round(cumsum(three_fg) / cumsum(three_fga), 3) * 100)
```


```{r, include=FALSE}
colnames(df22)
df22 <- df22[, c(1,2,3,5,4)]

df22 <- df22 %>% rename(
  three_fg = three_makes,
  three_fga = three_attempts,
  three_pct = pct
)

```


```{r, include=FALSE}
df21$season <- "2020-21"
df22$season <- "2021-22"

dfall <- rbind(df21, df22)
```



```{r, include=FALSE}
#subset(dfall, game>= 25 & season == "2020-21")

three <- ggplot(dfall, aes(x=game, y=running_pct, group=season), size = 1.5) +
  geom_line(aes(color=season), size= 1) + 
  theme_minimal() + 
  labs(y = "Running 3pt Percentage", 
       x= "Game")

three <- three + scale_color_manual(breaks = c("2020-21", "2021-22"),
                        values=c("orange", "black"))
```


# Season Comparison

```{r, echo=FALSE, dpi = 200}
three <- three + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

print(three)
```


